TEMPERATURES AND METALLICITIES OF M DWARFS IN THE APOGEE SURVEY

### 2020-01-16 paper/paper.tex
add submitted version

\received{\color{red} MM/DD/YY}
\revised{\color{red} MM/DD/YY}
\accepted{\color{red} MM/DD/YY}
\submitjournal{\apj}

\shorttitle{temperatures and metallicities for m dwarfs}

@@

@@

Following the procedure of \citet{Ness:2015} and \citet{Ho:2017a} we adopt a simple linear model that assumes the flux at each pixel of the spectrum can be parameterized as a function of a label vector $\ell$ and coefficient vector \emph{$\theta$}. For each star \emph{n}, at wavelength pixel \emph{$\lambda$}, we assume that the measured flux for a star at a given pixel is the sum of the coefficient and label product, and observational noise:
\begin{equation}
	f_{n\lambda} = \theta_{\lambda}^{T} \cdot \ell_{n} + N_{\lambda} \label{eqn:cannon}
\end{equation} 
\color{red}{\bf Here we use the noise model }\color{black} $N_{\lambda}=[s_{\lambda}^2 + \sigma_{n\lambda}^2]\xi_{n\lambda}$, where the bracketed term is the root mean sum of the intrinsic scatter of the model at each pixel \emph{$s_{\lambda}$}, and the uncertainty due to instrumental effects \emph{$\sigma_{n\lambda}$}, which is then multiplied times a Gaussian random number $\xi_{n\lambda} \sim \mathcal{N} (0,1)$. Equation (\ref{eqn:cannon}) corresponds to the single-pixel log-likelihood function:
\begin{equation}
	\ln p(f_{n\lambda}|\theta^T_{\lambda}, \ell_n, s^2_n) = 
	-\frac{1}{2}\frac{[f_{n\lambda} - \theta_{\lambda}^{T} \cdot \ell_{n}]^2}{s_{\lambda}^2 + \sigma_{n\lambda}^2} - \frac{1}{2} \ln(s_{\lambda}^2 + \sigma_{n\lambda}^2) 
	\label{eqn:likelihood}
\end{equation}
which gives the probability density function of the measured flux, given the labels, coefficients and scatters.

@@

\thecannon\ model can in principle be trained on any physical or empirical labels available \color{red}{\bf beyond those which typically parameterize theoretical atmospheric models ($\teff$, logg, [M/H], etc.), such as additional physical parameters (e.g., mass/age \citealt{Ness:2016}) or empirical proxies for physical parameters (e.g., spectral types, colors, magnitudes), }\color{black} giving a wide range of flexibility to the model. 
However, choosing a training sample with high quality labels is critical to its performance.
Limitations of \thecannon\ include that test (output) labels are only accurate if the training labels are accurate, and only precise if the training labels are measured consistently across the training sample.
It is also critical to have a training sample with the dynamic range to span the entire parameter space of interest, as \thecannon\ does not extrapolate well outside the parameter space of the training sample.
Finally, \thecannon\ assumes that the dependence of \color{red}{\bf the }\color{black} spectrum on labels is continuous and smooth---and in this implementation, well approximated by quadratic functions. If that isn't true, there will be features that \thecannon\ cannot reproduce.

For the purpose of this study, we have constructed two different training samples: first a one-dimensional \emph{spectral type model}, and second, a two-dimensional \emph{physical parameter model}, which describes the temperature and metallicity. The choice of training labels, dimensionality of our data set, and requirements for a good training set \color{red}{\bf are discussed }\color{black} further in Section \ref{sec:discussion}.

@@

The physical parameter training sample consists of 87 sources with reference labels distributed over $2859 <T_{\rm eff} < 4131$K, and $-0.48 < \feh < 0.49$\,dex, 41 sources of which are drawn from \citet{Mann:2015} (hereafter M15), and 46 of which are part of a previously unpublished extension sample to M15, analyzed using similar data and identical techniques to M15. The major difference between the extension sample is that the sample had lower-quality or no parallaxes (prior to Gaia data) and hence were omitted from the M15 study and were less vetted for binarity than the M15 sample \color{red}{\bf (however, all sources in the training sample were visually inspected by color-magnitude position for binarity before addition).}\color{black}

@@

\begin{figure}
	\includegraphics[width=\linewidth]{crv_test_teff.png}
	\includegraphics[width=\linewidth]{crv_test_fe_h.png}
	\caption{Consistency test for the Mann-trained temperature/metallicity type model; self-test (top), and leave-one-out cross validation (bottom). Sources with strong rotational broadening (2M08155393+3136392, 2M09174473+4612246, 2M10331367+3409120, 2M15594729+4403595, 2M13400879+4346380), as identified by visual inspection are marked with red circles. Predictive accuracy, as computed by the scatter in cross validation is 77K in temperature, and 0.09 dex in metallicity.} 
	\label{fig:mann_validation}
\end{figure}

@@

Another mode of analysis we can \color{red}{\bf utilize }\color{black} with \thecannon\ is how the derivative of the model changes with respect to given training parameters, which makes our model interpretable for discovering or verifying atomic or molecular lines with strong dependence on different physical parameters. The top two panels of Figures \ref{fig:demo_teff} and \ref{fig:demo_feh} show two example spectra and model fits for two different temperatures (Figure \ref{fig:demo_teff}), and two different metallicities (Figure \ref{fig:demo_feh}), with atomic and molecular features identified by the abundance analysis of \citet{Souto:2017}. The bottom panels of Figures \ref{fig:demo_teff} and \ref{fig:demo_feh} show the derivative of flux with respect to temperature and metallicity at each pixel, taken at the median training values. In order to evaluate which spectral features show statistically significant change with respect to input label, we compute \color{red}{\bf the }\color{black} error of the derivative at each pixel using a jackknife statistic (with a 1-$\sigma$ level overplotted in red):
\begin{equation}
	\sigma_{\theta,m}^2 = \frac{N-1}{N} \sum^N_{n=1} (\theta_{/n}^m - \theta^m)^2 
\end{equation}
where $\sigma_{\theta,m}$ is the error at pixel \emph{m}, \emph{N} is the total number of stars in the sample indexed by \emph{n}, $\theta$ is the coeffient vector trained on all \emph{N} sources, and $\theta_{/n}$ is the coefficient vector trained on \emph{N-1} sources excluding star \emph{n}. A summary of identified lines with derivative values greater than $2\sigma_{\rm jackkife}$ is given in Table \ref{table:teff_derivative} and Table \ref{table:feh_derivative}.

@@

The spectra contain roughly 8000 pixels,
so we might expect the $\chi^2$ values to be close to 8000 in magnitude,
\color{red}{\bf however, they }\color{black} are much higher.
This discrepancy flows from the fact that, while the spectral model is good at the percent
level, the signal-to-noise of the typical spectra are more than 100.
That is, the $\chi^2$ values do show that the model is not good in the frequentist
sense; it is only good at the percent level.

\begin{figure}
	\begin{center}
	\includegraphics[width=\linewidth]{crv_test_spt.png}
	\end{center}
	\caption{Leave-one-out cross validation test for the West-trained spectral type model. Predictive accuracy, as computed by the scatter in cross validation is 0.9 subtypes.}
	\label{fig:west_validation}
\end{figure}

@@

We trained \thecannon\ on 51 M dwarfs ranging \color{red}{\bf between }\color{black} M0$-$M9 with a one-dimensional spectral type label, and obtained a precision of $\pm$0.9 spectral types, also similar to the original training label uncertainty of $\pm$1 spectral type. We note however, that the training sample is distributed heavily towards earlier type sources, with a median spectral type of 3 and only one M8 and one M9 source. As seen in Figure \ref{fig:west_validation}, the model performs poorly at reproducing spectral types $>$M8, which confirms that \thecannon\ does not extrapolate well to labels outside of the training sample space. Because of this skew for late-type sources, we report our spectral type model to be precise to $\pm$0.7 spectral types for the range M0-M6. Repeating the analysis of Section \ref{subsec:mann_results}, Figure \ref{fig:west_validation} shows LOOCV test for the labels reported in Table \ref{table:west_results}, and Figure \ref{fig:west_derivative} shows the derivative of model flux with varying spectral type.

@@

\begin{figure*}
	\includegraphics[width=\linewidth]{cannon_chi_dist.png}
	\caption{\textit{Left:} the distribution of fits for the training sample and known samples of pre-MS stars (\citealt{Cottaar:2014}) and binary sources \citep{ElBadry:2018,Skinner:2018}; \textit{right:} the distribution of $\chi^2$ fits for all 14,827 sources in the APOGEE--\gaia\ cross match, with color cuts $1<G_{BP}-G_{RP}<6$ and $7.5<M_{G}<20$ and $\varpi>0$. We apply a quality cut of $\chi^2 < 100,000$ to the test sample for those sources we report as ``safe''.}
	\label{fig:chi_dist}
\end{figure*}

\begin{figure*}
	\includegraphics[width=\linewidth]{cmd_selection.png}
	\caption{\gaia\ and 2MASS color-magnitude cuts for the 12,037 sources with $\chi^2<100,000$. Over plotted with orange triangles are the 67 out of 87 training sample sources which have parallaxes measured by \gaia. The coordinates for the selected quadrangles are
	$\left\{ (1.4, 7.5), (2.2, 7.5), (4.2, 14), (3.3, 14) \right\}$ corresponding to ($BP-RP$, $M_G$) for the \gaia\ color-magnitudes shown in the left panel, and 
	$\left\{ (.7, 4.3), (1.1, 4.3), (.7, 9), (1.1, 9) \right\}$ corresponding to ($J-K$, $M_K$) for the 2MASS color-magnitudes shown in the right panel. }
	\label{fig:cmd_selection}
\end{figure*}

@@

\begin{enumerate}
\item \textbf{Quality of fit cut:} We apply a \cannon\ model $\chi^2$ cut of less than 100,000, chosen to remove badly fit sources (such as fast rotators), but include $\chi^2$ values close \color{red}{\bf to }\color{black} the distribution of training sample (Figure \ref{fig:chi_dist}).

\item \textbf{Color-magnitude cuts:} Using \gaia\ and 2MASS photometry we apply the \color{red}{\bf additional }\color{black} color-magnitude selections shown in Figure \ref{fig:cmd_selection} to remove sources above the main sequence (which are likely pre-main sequence, reddened K dwarfs and\color{red}{\bf /or }\color{black} multiples), and subdwarfs below the main sequence.

@@

\begin{figure}
\begin{center}
	\includegraphics[width=\linewidth]{ruwe_cut_before.png} \\
	\includegraphics[width=\linewidth]{ruwe_cut_after.png}
	\caption{The left panel shows the test sample of 8,335 sources after applying Selections $1-3$ described in Section \ref{subsec:test_selection}. The right panel shows the same test sample reduced to 6,221 after applying an astrometric quality cut of ${\rm RUWE}<1.2$ (Selection 4). To further remove sources which were likely binary contamination (Selection 5), we cut out sources above the red line which sparsely lay above the majority of the main-sequence with temperatures and metallicities that deviate from the expected gradient, reducing the final test sample to a size of 5,875 sources. The red line shown is constrained by the ($BP-RP$, $M_G$) coordinates $\left\{ (3.5,12), (1.87,7.5) \right\}$.}
	\label{fig:ruwe_cut}
\end{center}
\end{figure}

\begin{figure*}
\begin{center}
	\includegraphics[width=.9\linewidth]{aspcap_cannon_validation.png}
	\caption{\cannon\ (right column) and ASPCAP (left column) temperatures as compared to temperatures estimated color-temperature relations from three different literature sources for a subsample of 1,702 sources with $V,J,K$ colors. Shown are \cannon\ and ASPCAP temperatures (x-axis) compared to temperatures estimated from the \citealt{Mann:2015} $V-J$ and $\feh$ dependent relation (top), and the \citealt{Boyajian:2012} relation $V-K$ and $\feh$ (bottom).}
	\label{fig:teff_comparisons}
\end{center}
\end{figure*}

\begin{figure}
\begin{center}
	\includegraphics[width=\linewidth]{wide_binary_met.png}
	\caption{Metallicities of \cannon\ M dwarfs in binaries (x-axis) compared to the metallicities of their companions (y-axis). Label colors show the source and spectral type of the companion metallicities. Metallicity values for each binary pair is compiled in Table \ref{table:binaries}. The overall scatter between the 15 metallicity pairs is 0.08\,dex.} 
	\label{fig:wide_binary_met}
\end{center}
\end{figure}

@@

\begin{figure*}
\begin{center}
	\includegraphics[height=5cm]{toomre.png}
	\includegraphics[height=5cm]{met_hist.png}
	\caption{Toomre diagram showing the distribution of the \cannon\ metallicities (left). Dashed lines at $v_{tot}=70$\,km/s and 180\,km/s show roughly the separation of thin/thick/halo stars. Histogram (right) shows the metallicity distribution of thin and thick disk stars, with blue corresponding to $v_{tot}<70$\,km/s and green corresponding to $70<v_{tot}<180$\,km/s respectively. The average/standard deviation metallicity of the thin disk distribution is $\feh=0.00\pm0.17\,$dex, and the average/standard deviation metallicity of the thick disk distribution is $\feh=-0.14\pm0.19\,$dex.}
	\label{fig:toomre}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
	\includegraphics[width=.3\linewidth]{met_comp1.png}
	\includegraphics[width=.3\linewidth]{met_comp2.png}
	\includegraphics[width=.3\linewidth]{met_comp3.png} 
	\includegraphics[width=.3\linewidth]{met_comp4.png}
	\includegraphics[width=.3\linewidth]{met_comp5.png}
	\includegraphics[width=.3\linewidth]{met_comp6.png} 
	\includegraphics[width=.3\linewidth]{met_comp7.png}
	\includegraphics[width=.3\linewidth]{met_comp8.png}
	\includegraphics[width=.3\linewidth]{met_comp9.png} 
	\caption{A comparison of ASPCAP and Cannon metallicities for 8,335 test sample sources separated into temperature bins of 100K. We see the scatter and metal-poor bias of ASPCAP metallicities clearly increases at decreasing temperatures.} 
	\label{fig:met_comp}
\end{center}
\end{figure*}

@@

As a check of our test sample metallicity reliability, we cross-matched our M dwarf final sample with the \citet{Elbadry:2018b} catalog of $>$50,000 high-confidence, widely separated binaries identified by \drtwo. 
In total we found 216 of the APOGEE M dwarfs to have binary pairs (46 FGK+M, 155 M+M, and 15 WD+M). Out of the 155 M+M pairs, 8 pairs contained both pairs in APOGEE. Cross-matching the list of FGK+M dwarf companions with several catalogs/surveys with measured stellar metallicities, we found an additional 7 sources with FGK metallicities from LAMOST \citep{Zhao:2012} and APOGEE (ASPCAP). The metallicity measurements for the 15 M dwarf binaries and their companions are given in Table \ref{table:binaries} \color{red}{\bf and shown in Figure \ref{fig:wide_binary_met} }\color{black}, and the overall scatter is $0.08\,$dex--an improvement over the scatter of ASPCAP metallicities which is $0.15\,$dex for these 15 sources. Internal consistency of the two models (the scatter of the 8 M+M pairs both in APOGEE) are 0.06\,dex and 0.12\,dex for \cannon\ and ASPCAP respectively.

@@

\begin{figure*}
\begin{center}
	\includegraphics[width=.47\linewidth]{cmd_teff_full.png}
	\includegraphics[width=.47\linewidth]{cmd_teff_safe.png} \\
	\includegraphics[width=.47\linewidth]{cmd_feh_full.png}
	\includegraphics[width=.47\linewidth]{cmd_feh_safe.png} \\
	\includegraphics[width=.47\linewidth]{cmd_spt_full.png}
	\includegraphics[width=.47\linewidth]{cmd_spt_safe.png}
	\caption{Full sample of 14,827 M dwarfs colored by \cannon\ labels before selection (left) and final sample selection of 5,875 M dwarfs after applying selection criteria described in Section \ref{subsec:test_selection} (right), to reduce contamination from sources with that are not similar to the training sample (not single, main-sequence M stars, such as pre-main sequence, spectroscopic binaries, and K dwarfs). Overplotted with orange triangles are the M15 and W11 training samples, for their respective \cannon\ test labels. Temperature gradient increases with decreasing color, spectral subtype increases with increasing color, and metallicity gradient increases perpendicularly up from the main sequence branch as expected. Deviations from these gradients seen at the upper boundary of the main sequence is likely remaining contamination from the binary sequence.}
	\label{fig:safe_selection}
\end{center}
\end{figure*}

@@

While a detailed analysis of atmospheric model limitations is beyond the scope of this paper, 
our results provide an avenue to compare the FGK star metallicity scale 
to the less well-understood M dwarf metallicity scale.
These results find that atmospheric metallicities are systematically metal-poor biased
compared to \cannon-based metallicities trained on sources with metallicities calibrated
to those of FGK companions. 
At the high temperature end ($\teff > 3600$\,K), the ASPCAP metallicity bias is $-0.12-0.15\,$dex with a scatter of $0.05-0.06\,$dex relative to \cannon\ metallicities, and increases to a bias of $-0.3\,$dex and scatter of $0.13\,$dex at the lower temperature end ($\teff < 3200$\,K) (Figure \ref{fig:met_comp}).
We suspect that this metal-poor bias, while not explored to a great extent in this work, is due 
to the line lists of the models: 
an effect in which the optimizer of the pipeline may be lowering the continuum level and metallicity of the fit to compensate for the missing lines or opacities.
\color{red}{\bf We also note that this analysis was completed using data from Data Release 14 of APOGEE, 
which did include molecular lines from FeH in the pipeline at the time, which become numerous and strong for $\teff \lesssim 3600$\,K. }\color{black}
Further analysis would need to be done to \color{red}{\bf quantify the metallicity improvement for M dwarfs in later data releases of APOGEE now available (in which FeH has been included in analysis)}\color{black}, and determine if the metallicity bias
is found in other model grids (besides the ATLAS/MARCS models used by the ASPCAP pipeline), 
and if the effect is present at other wavelengths.

>>>>

\submitjournal{\apj}

\shorttitle{temperatures and metallicities for m dwarfs}
\shortauthors{birky et al.}

@@

\correspondingauthor{Jessica Birky}
\email{jbirky@uw.edu}

\author[0000-0002-7961-6881]{Jessica Birky}
\affil{Center for Astrophysics and Space Science, University of California San Diego, La Jolla, CA 92093, USA}
\affil{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}

\author[0000-0003-2866-9403]{David W. Hogg}
\affil{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726
Broadway, New York, NY 10003, USA}
\affil{Center for Data Science, New York University, 60 Fifth Ave, New York, NY 10011, USA}
\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}

\author[0000-0003-3654-1602]{Andrew W. Mann}
\affil{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY 10027, USA}
\affil{Department of Physics and Astronomy, The University of North Carolina at Chapel Hill, Chapel Hill, NC 27599, USA}

\author[0000-0002-6523-9536]{Adam Burgasser}
\affil{Center for Astrophysics and Space Science, University of California San Diego, La Jolla, CA 92093, USA}

@@

Following the procedure of \citet{Ness:2015} and \citet{Ho:2017a} we adopt a simple linear model that assumes the flux at each pixel of the spectrum can be parameterized as a function of a label vector $\ell$ and coefficient vector \emph{$\theta$}. For each star \emph{n}, at wavelength pixel \emph{$\lambda$}, we assume that the measured flux for a star at a given pixel is the sum of the coefficient and label product, and observational noise:
\begin{equation}
	f_{n\lambda} = \theta_{\lambda}^{T} \cdot \ell_{n} + N_{\lambda} \label{eqn:cannon}
\end{equation} 
Here we use the noise model $N_{\lambda}=[s_{\lambda}^2 + \sigma_{n\lambda}^2]\xi_{n\lambda}$, where the bracketed term is the root mean sum of the intrinsic scatter of the model at each pixel \emph{$s_{\lambda}$}, and the uncertainty due to instrumental effects \emph{$\sigma_{n\lambda}$}, which is then multiplied times a Gaussian random number $\xi_{n\lambda} \sim \mathcal{N} (0,1)$. Equation (\ref{eqn:cannon}) corresponds to the single-pixel log-likelihood function:
\begin{equation}
	\ln p(f_{n\lambda}|\theta^T_{\lambda}, \ell_n, s^2_n) = 
	-\frac{1}{2}\frac{[f_{n\lambda} - \theta_{\lambda}^{T} \cdot \ell_{n}]^2}{s_{\lambda}^2 + \sigma_{n\lambda}^2} - \frac{1}{2} \ln(s_{\lambda}^2 + \sigma_{n\lambda}^2) 
	\label{eqn:likelihood}
\end{equation}
which gives the probability density function of the measured flux, given the labels, coefficients and scatters.

@@

\thecannon\ model can in principle be trained on any physical or empirical labels available beyond those which typically parameterize theoretical atmospheric models ($\teff$, logg, [M/H], etc.), such as additional physical parameters (e.g., mass/age \citealt{Ho:2017b}) or empirical proxies for physical parameters (e.g., spectral types, colors, magnitudes), giving a wide range of flexibility to the model. 
However, choosing a training sample with high quality labels is critical to its performance.
Limitations of \thecannon\ include that test (output) labels are only accurate if the training labels are accurate, and only precise if the training labels are measured consistently across the training sample.
It is also critical to have a training sample with the dynamic range to span the entire parameter space of interest, as \thecannon\ does not extrapolate well outside the parameter space of the training sample.
Finally, \thecannon\ assumes that the dependence of the spectrum on labels is continuous and smooth---and in this implementation, well approximated by quadratic functions. If that isn't true, there will be features that \thecannon\ cannot reproduce.

For the purpose of this study, we have constructed two different training samples: first a one-dimensional \emph{spectral type model}, and second, a two-dimensional \emph{physical parameter model}, which describes the temperature and metallicity. The choice of training labels, dimensionality of our data set, and requirements for a good training set are discussed further in Section \ref{sec:discussion}.

@@

The physical parameter training sample consists of 87 sources with reference labels distributed over $2859 <T_{\rm eff} < 4131$K, and $-0.48 < \feh < 0.49$\,dex, 41 sources of which are drawn from \citet{Mann:2015} (hereafter M15), and 46 of which are part of a previously unpublished extension sample to M15, analyzed using similar data and identical techniques to M15. The major difference between the extension sample is that the sample had lower-quality or no parallaxes (prior to Gaia data) and hence were omitted from the M15 study and were less vetted for binarity than the M15 sample (however, all sources in the training sample were visually inspected by color-magnitude position for binarity before addition).

@@

\begin{figure}
	\includegraphics[width=\linewidth]{figs/crv_test_teff.png}
	\includegraphics[width=\linewidth]{figs/crv_test_fe_h.png}
	\caption{Consistency test for the Mann-trained temperature/metallicity type model; self-test (top), and leave-one-out cross validation (bottom). Sources with strong rotational broadening (2M08155393+3136392, 2M09174473+4612246, 2M10331367+3409120, 2M15594729+4403595, 2M13400879+4346380), as identified by visual inspection are marked with red circles. Predictive accuracy, as computed by the scatter in cross validation is 77K in temperature, and 0.09 dex in metallicity.} 
	\label{fig:mann_validation}
\end{figure}

@@

Another mode of analysis we can utilize with \thecannon\ is how the derivative of the model changes with respect to given training parameters, which makes our model interpretable for discovering or verifying atomic or molecular lines with strong dependence on different physical parameters. The top two panels of Figures \ref{fig:demo_teff} and \ref{fig:demo_feh} show two example spectra and model fits for two different temperatures (Figure \ref{fig:demo_teff}), and two different metallicities (Figure \ref{fig:demo_feh}), with atomic and molecular features identified by the abundance analysis of \citet{Souto:2017}. The bottom panels of Figures \ref{fig:demo_teff} and \ref{fig:demo_feh} show the derivative of flux with respect to temperature and metallicity at each pixel, taken at the median training values. In order to evaluate which spectral features show statistically significant change with respect to input label, we compute the error of the derivative at each pixel using a jackknife statistic (with a 1-$\sigma$ level overplotted in red):
\begin{equation}
	\sigma_{\theta,m}^2 = \frac{N-1}{N} \sum^N_{n=1} (\theta_{/n}^m - \theta^m)^2 
\end{equation}
where $\sigma_{\theta,m}$ is the error at pixel \emph{m}, \emph{N} is the total number of stars in the sample indexed by \emph{n}, $\theta$ is the coeffient vector trained on all \emph{N} sources, and $\theta_{/n}$ is the coefficient vector trained on \emph{N-1} sources excluding star \emph{n}. A summary of identified lines with derivative values greater than $2\sigma_{\rm jackkife}$ is given in Table \ref{table:teff_derivative} and Table \ref{table:feh_derivative}.

@@

The spectra contain roughly 8000 pixels,
so we might expect the $\chi^2$ values to be close to 8000 in magnitude,
however, they are much higher.
This discrepancy flows from the fact that, while the spectral model is good at the percent
level, the signal-to-noise of the typical spectra are more than 100.
That is, the $\chi^2$ values do show that the model is not good in the frequentist
sense; it is only good at the percent level.

\begin{figure}
	\begin{center}
	\includegraphics[width=\linewidth]{figs/crv_test_spt.png}
	\end{center}
	\caption{Leave-one-out cross validation test for the West-trained spectral type model. Predictive accuracy, as computed by the scatter in cross validation is 0.9 subtypes.}
	\label{fig:west_validation}
\end{figure}

@@

We trained \thecannon\ on 51 M dwarfs ranging between M0$-$M9 with a one-dimensional spectral type label, and obtained a precision of $\pm$0.9 spectral types, also similar to the original training label uncertainty of $\pm$1 spectral type. We note however, that the training sample is distributed heavily towards earlier type sources, with a median spectral type of 3 and only one M8 and one M9 source. As seen in Figure \ref{fig:west_validation}, the model performs poorly at reproducing spectral types $>$M8, which confirms that \thecannon\ does not extrapolate well to labels outside of the training sample space. Because of this skew for late-type sources, we report our spectral type model to be precise to $\pm$0.7 spectral types for the range M0-M6. Repeating the analysis of Section \ref{subsec:mann_results}, Figure \ref{fig:west_validation} shows LOOCV test for the labels reported in Table \ref{table:west_results}, and Figure \ref{fig:west_derivative} shows the derivative of model flux with varying spectral type.

@@

\begin{figure*}
	\includegraphics[width=\linewidth]{figs/cannon_chi_dist.png}
	\caption{\textit{Left:} the distribution of fits for the training sample and known samples of pre-MS stars (\citealt{Cottaar:2014}) and binary sources \citep{ElBadry:2018,Skinner:2018}; \textit{right:} the distribution of $\chi^2$ fits for all 14,827 sources in the APOGEE--\gaia\ cross match, with color cuts $1<G_{BP}-G_{RP}<6$ and $7.5<M_{G}<20$ and $\varpi>0$. We apply a quality cut of $\chi^2 < 100,000$ to the test sample for those sources we report as ``safe''.}
	\label{fig:chi_dist}
\end{figure*}

\begin{figure*}
	\includegraphics[width=\linewidth]{figs/cmd_selection.png}
	\caption{\gaia\ and 2MASS color-magnitude cuts for the 12,037 sources with $\chi^2<100,000$. Over plotted with orange triangles are the 67 out of 87 training sample sources which have parallaxes measured by \gaia. The coordinates for the selected quadrangles are
	$\left\{ (1.4, 7.5), (2.2, 7.5), (4.2, 14), (3.3, 14) \right\}$ corresponding to ($BP-RP$, $M_G$) for the \gaia\ color-magnitudes shown in the left panel, and 
	$\left\{ (.7, 4.3), (1.1, 4.3), (.7, 9), (1.1, 9) \right\}$ corresponding to ($J-K$, $M_K$) for the 2MASS color-magnitudes shown in the right panel. }
	\label{fig:cmd_selection}
\end{figure*}

@@

\begin{enumerate}
\item \textbf{Quality of fit cut:} We apply a \cannon\ model $\chi^2$ cut of less than 100,000, chosen to remove badly fit sources (such as fast rotators), but include $\chi^2$ values close to the distribution of training sample (Figure \ref{fig:chi_dist}).

\item \textbf{Color-magnitude cuts:} Using \gaia\ and 2MASS photometry we apply the additional color-magnitude selections shown in Figure \ref{fig:cmd_selection} to remove sources above the main sequence (which are likely pre-main sequence, reddened K dwarfs and/or multiples), and subdwarfs below the main sequence.

@@

\begin{figure}
\begin{center}
	\includegraphics[width=\linewidth]{figs/ruwe_cut_before.png} \\
	\includegraphics[width=\linewidth]{figs/ruwe_cut_after.png}
	\caption{The left panel shows the test sample of 8,335 sources after applying Selections $1-3$ described in Section \ref{subsec:test_selection}. The right panel shows the same test sample reduced to 6,221 after applying an astrometric quality cut of ${\rm RUWE}<1.2$ (Selection 4). To further remove sources which were likely binary contamination (Selection 5), we cut out sources above the red line which sparsely lay above the majority of the main-sequence with temperatures and metallicities that deviate from the expected gradient, reducing the final test sample to a size of 5,875 sources. The red line shown is constrained by the ($BP-RP$, $M_G$) coordinates $\left\{ (3.5,12), (1.87,7.5) \right\}$.}
	\label{fig:ruwe_cut}
\end{center}
\end{figure}

\begin{figure*}
\begin{center}
	\includegraphics[width=.9\linewidth]{figs/aspcap_cannon_validation.png}
	\caption{\cannon\ (right column) and ASPCAP (left column) temperatures as compared to temperatures estimated color-temperature relations from three different literature sources for a subsample of 1,702 sources with $V,J,K$ colors. Shown are \cannon\ and ASPCAP temperatures (x-axis) compared to temperatures estimated from the \citealt{Mann:2015} $V-J$ and $\feh$ dependent relation (top), and the \citealt{Boyajian:2012} relation $V-K$ and $\feh$ (bottom).}
	\label{fig:teff_comparisons}
\end{center}
\end{figure*}

\begin{figure}
\begin{center}
	\includegraphics[width=\linewidth]{figs/wide_binary_met.png}
	\caption{Metallicities of \cannon\ M dwarfs in binaries (x-axis) compared to the metallicities of their companions (y-axis). Label colors show the source and spectral type of the companion metallicities. Metallicity values for each binary pair is compiled in Table \ref{table:binaries}. The overall scatter between the 15 metallicity pairs is 0.08\,dex.} 
	\label{fig:wide_binary_met}
\end{center}
\end{figure}

@@

\begin{figure*}
\begin{center}
	\includegraphics[height=5cm]{figs/toomre.png}
	\includegraphics[height=5cm]{figs/met_hist.png}
	\caption{Toomre diagram showing the distribution of the \cannon\ metallicities (left). Dashed lines at $v_{tot}=70$\,km/s and 180\,km/s show roughly the separation of thin/thick/halo stars. Histogram (right) shows the metallicity distribution of thin and thick disk stars, with blue corresponding to $v_{tot}<70$\,km/s and green corresponding to $70<v_{tot}<180$\,km/s respectively. The average/standard deviation metallicity of the thin disk distribution is $\feh=0.00\pm0.17\,$dex, and the average/standard deviation metallicity of the thick disk distribution is $\feh=-0.14\pm0.19\,$dex.}
	\label{fig:toomre}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
	\includegraphics[width=.3\linewidth]{figs/met_comp1.png}
	\includegraphics[width=.3\linewidth]{figs/met_comp2.png}
	\includegraphics[width=.3\linewidth]{figs/met_comp3.png} 
	\includegraphics[width=.3\linewidth]{figs/met_comp4.png}
	\includegraphics[width=.3\linewidth]{figs/met_comp5.png}
	\includegraphics[width=.3\linewidth]{figs/met_comp6.png} 
	\includegraphics[width=.3\linewidth]{figs/met_comp7.png}
	\includegraphics[width=.3\linewidth]{figs/met_comp8.png}
	\includegraphics[width=.3\linewidth]{figs/met_comp9.png} 
	\caption{A comparison of ASPCAP and Cannon metallicities for 8,335 test sample sources separated into temperature bins of 100K. We see the scatter and metal-poor bias of ASPCAP metallicities clearly increases at decreasing temperatures.} 
	\label{fig:met_comp}
\end{center}
\end{figure*}

@@

As a check of our test sample metallicity reliability, we cross-matched our M dwarf final sample with the \citet{Elbadry:2018b} catalog of $>$50,000 high-confidence, widely separated binaries identified by \drtwo. 
In total we found 216 of the APOGEE M dwarfs to have binary pairs (46 FGK+M, 155 M+M, and 15 WD+M). Out of the 155 M+M pairs, 8 pairs contained both pairs in APOGEE. Cross-matching the list of FGK+M dwarf companions with several catalogs/surveys with measured stellar metallicities, we found an additional 7 sources with FGK metallicities from LAMOST \citep{Zhao:2012} and APOGEE (ASPCAP). The metallicity measurements for the 15 M dwarf binaries and their companions are given in Table \ref{table:binaries} and shown in Figure \ref{fig:wide_binary_met}, and the overall scatter is $0.08\,$dex--an improvement over the scatter of ASPCAP metallicities which is $0.15\,$dex for these 15 sources. Internal consistency of the two models (the scatter of the 8 M+M pairs both in APOGEE) are 0.06\,dex and 0.12\,dex for \cannon\ and ASPCAP respectively.

@@

\begin{figure*}
\begin{center}
	\includegraphics[width=.47\linewidth]{figs/cmd_teff_full.png}
	\includegraphics[width=.47\linewidth]{figs/cmd_teff_safe.png} \\
	\includegraphics[width=.47\linewidth]{figs/cmd_feh_full.png}
	\includegraphics[width=.47\linewidth]{figs/cmd_feh_safe.png} \\
	\includegraphics[width=.47\linewidth]{figs/cmd_spt_full.png}
	\includegraphics[width=.47\linewidth]{figs/cmd_spt_safe.png}
	\caption{Full sample of 14,827 M dwarfs colored by \cannon\ labels before selection (left) and final sample selection of 5,875 M dwarfs after applying selection criteria described in Section \ref{subsec:test_selection} (right), to reduce contamination from sources with that are not similar to the training sample (not single, main-sequence M stars, such as pre-main sequence, spectroscopic binaries, and K dwarfs). Overplotted with orange triangles are the M15 and W11 training samples, for their respective \cannon\ test labels. Temperature gradient increases with decreasing color, spectral subtype increases with increasing color, and metallicity gradient increases perpendicularly up from the main sequence branch as expected. Deviations from these gradients seen at the upper boundary of the main sequence is likely remaining contamination from the binary sequence.}
	\label{fig:safe_selection}
\end{center}
\end{figure*}

@@

While a detailed analysis of atmospheric model limitations is beyond the scope of this paper, 
our results provide an avenue to compare the FGK star metallicity scale 
to the less well-understood M dwarf metallicity scale.
These results find that atmospheric metallicities are systematically metal-poor biased
compared to \cannon-based metallicities trained on sources with metallicities calibrated
to those of FGK companions. 
At the high temperature end ($\teff > 3600$\,K), the ASPCAP metallicity bias is $-0.12-0.15\,$dex with a scatter of $0.05-0.06\,$dex relative to \cannon\ metallicities, and increases to a bias of $-0.3\,$dex and scatter of $0.13\,$dex at the lower temperature end ($\teff < 3200$\,K) (Figure \ref{fig:met_comp}).
We suspect that this metal-poor bias, while not explored to a great extent in this work, is due 
to the line lists of the models: 
an effect in which the optimizer of the pipeline may be lowering the continuum level and metallicity of the fit to compensate for the missing lines or opacities.
We also note that this analysis was completed using data from Data Release 14 of APOGEE, 
which did include molecular lines from FeH in the pipeline at the time, which become numerous and strong for $\teff \lesssim 3600$\,K \citep{Souto:2017}.
Further analysis would need to be done to quantify the metallicity improvement for M dwarfs in future data releases of APOGEE, and determine if the metallicity bias
is found in other model grids (besides the ATLAS/MARCS models used by the ASPCAP pipeline), 
and if the effect is present at other wavelengths.
